Target Speaker Extraction using Discrete Representations from Self-Supervised Models and Language Models \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-09-04 src/paper.tex
update

Discretization methods have been studied for speech 
enhancement \cite{selm} and blind speech separation \cite{dasb,tokensplit}, 
however, this approach has 
been rarely studied for target speaker extraction. In this paper, we present a novel way to do 
target speaker extraction using discrete tokens and LMs. Inspired by the blind speech separation 
networks proposed in DASB \cite{dasb}, our model has three stages: encoding, modeling and decoding.
For the encoding stage, reference speech and mixed speech are tokenized using WavLM and 
Kmeans as tokenizer. Unlike SELM \cite{selm}, which uses the 6th layer of wavlm as input to the Kmeans layer, 
we follow the recipes in \cite{dasb} and uses the output from hidden 
layers 1, 3, 7, 12, 18, 23 of WavLM as input to the Kmeans layer. 6 Kmeans models
are trained and applied on each layer output respectively. 
Our results indicate that using multiple layers as 
input is better 
than using only the 6th layer. The reference speech is fed to the encoder directly. 
However, for the mixed speech, instead of directly passing it to the WavLM model, we first concatenated the reference speech on both sides 
of it before passing. After tokenization, we use only the tokens 
corresponding to the mixed speech. For the modeling stage, we utilized an attention 
embedding mechanism to add each embeddings of all layers together. We used a Cross 
Attention mechanism similar in \cite{sef_net} to inject the speaker information. 
An encoder-only LM is applied to

For the decoding stage, we used the pretrained unit HiFi-GAN 
in \cite{unit_hifi} to reconstruct the 
discrete tokens back to audio. Unlike SELM \cite{selm}, where a conformer 
detokenizer is trained 
to reconstruct the WavLM embeddings using Kmeans center embeddings before HiFi-GAN, \cite{unit_hifi} has proposed a scalable HiFi-GAN using dropout mechanism to directly 
reconstruct audio from multiple layers of tokens without a conformer detokenizer. It 
has also eliminated the trouble of training a HiFi-GAN for each layer. The encoder and 
decoder is freezed during the training.

>>>>

Discretization methods have been studied for speech 
enhancement \cite{selm} and blind speech separation \cite{dasb,tokensplit}, 
however, this approach has 
been rarely studied for target speaker extraction. In this paper, we present a novel way to do 
target speaker extraction using discrete tokens and LMs. Inspired by the blind speech separation 
networks proposed in DASB \cite{dasb}, our model has three stages: encoding, modeling and decoding.
For the encoding stage, reference speech and mixed speech are tokenized using WavLM and 
Kmeans as tokenizer. Unlike SELM \cite{selm}, which uses the 6th layer of WavLM as input to the Kmeans layer, 
we follow the recipes in \cite{dasb} and uses the output from hidden 
layers 1, 3, 7, 12, 18, 23 of WavLM as input to the Kmeans layer. 6 Kmeans models
are trained and applied on each layer output respectively. 
Our results indicate that using multiple layers as 
input is better 
than using only the 6th layer. The reference speech is fed to the encoder directly. 
However, for the mixed speech, instead of directly passing it to the WavLM model, we first concatenated the reference speech on both sides 
of it before passing. After tokenization, we use only the tokens 
corresponding to the mixed speech. For the modeling stage, we utilized an attention 
embedding mechanism to add each embeddings of all layers together. We used a Cross 
Attention mechanism similar in \cite{sef_net} to inject the speaker information. 
An encoder-only LM, following by a linear classifier, is applied after the cross attention module
to output the reconstructed tokens. 
For the decoding stage, we used the pretrained unit HiFi-GAN 
in \cite{unit_hifi} to reconstruct the 
discrete tokens back to audio. Unlike SELM \cite{selm}, where a conformer 
detokenizer is trained 
to reconstruct the WavLM embeddings using Kmeans center embeddings before HiFi-GAN, \cite{unit_hifi} has proposed a scalable HiFi-GAN using dropout mechanism to directly 
reconstruct audio from multiple layers of tokens without a conformer detokenizer. It 
has also eliminated the trouble of training a HiFi-GAN for each layer. The encoder and 
decoder is freezed during the training. The model overview is in Fig.\ref{model}. 
Through extensive studies, we have demonstrated this method has achieved comparable results in terms 
of speech quality and intelligibility. To the best of our knowledge, we are the first to explore 
using discrete tokens to conduct target speaker extraction.
